---
title: <center> <h1>Covid19 in Maranhão</h1> </center>
author: <center> <h2>Gabriel Simões Gaspar</h2> </center>
output: 
  html_document: 
    theme: united
    highlight: tango
    df_print: paged
---

This is a guide to plot an animated map with the evolution of cases of covid-19 for the Brazilian state of Maranhão. In order to do so, please follow the steps described below.
```{r include = FALSE}
rm(list=ls())
```

First begin by setting paths that will be used later on to locate the data and the shapefiles
```{r include = TRUE, warning = FALSE}
datapath <- "/Users/gabrielsgaspar/Desktop/Code/r/covid_ma/data"
shppath <- "/Users/gabrielsgaspar/Desktop/Code/r/covid_ma/shapefiles"
```

Now, if necessary, download and load the following libraries:
```{r include = TRUE, warning = FALSE, message = FALSE}
library(sf) 
library(tidyverse) 
library(gganimate)
```

## Organizing the dataset

We then beging by import the shapefiles for the state of Maranhão
```{r include = TRUE, warning = FALSE, message = FALSE}
# Shapefile for states
brasil1 <- st_read(dsn = "shapefiles/BRA_adm1.shp")
```

And now the municipalities in Maranhão
```{r include = TRUE, warning = FALSE, message = FALSE}
# Shapefile for municipalities
brasil2 <- st_read(dsn = "shapefiles/BRA_adm2.shp")
```

Since we are only working with the state of Maranhão we need to filter it in our dataset
```{r include = TRUE, warning = FALSE, message = FALSE}
# Filtering for Maranhão
ma_1 <- brasil1 %>% 
  filter(NAME_1 == "Maranhão")
```

Similarly, we filter the municipalities within Maranhão
```{r include = TRUE, warning = FALSE, message = FALSE}
# Filtering municipalities in Maranhão and renaming column
ma_2 <- brasil2 %>% 
  filter(NAME_1 == "Maranhão") %>%
  select(NAME_2) %>%
  rename(city = NAME_2)
  
# Calculate centroid of municipalites
ma_centroids <- st_centroid(ma_2)
```

Import the data for covid-19 cases in Maranhão from brasil.io 
```{r include = TRUE, warning = FALSE, message = FALSE}
# Downloading files
download.file(url = "https://brasil.io/dataset/covid19/caso_full/?state=MA&is_repeated=False&format=csv",
              destfile = "data/covid_ma")

# Importing files
covid_ma <- read_csv("data/covid_ma")
```

Perform join between shapefiles and data
```{r include = TRUE, warning = FALSE, message = FALSE}
# Left joining daframes
covid_ma <- left_join(ma_centroids, covid_ma)

# Droping N/A's
covid_ma <- covid_ma %>% 
  drop_na()
```

## Creating map

Now we can proceed with creating the map of covid-19 cases for the state of Maranhão and save it as a gif. In order to do so, execute the following code
```{r include = TRUE, warning = FALSE, message = FALSE}
# Make the base map
mapa_ma <- ggplot() +
  geom_sf(data = ma_1)

# Plot animation
animation_ma <- mapa_ma +
  geom_sf(data = covid_ma,
          aes(size = last_available_confirmed),
          colour = "blue", alpha = .5) +
  scale_size_continuous(breaks = c(1,10,100,500,1000,3000)) +
  labs(size = "Confirmed cases") +
  ggtitle("Evolution of covid-19 cases by municipality in Maranhão",
          subtitle = "Day {frame} of {nframes} since the first case") +
  labs(caption = "Source: brasil.io") +
  theme_minimal() +
  transition_manual(date)

# Setting speed of frames
animate(animation_ma, fps = 5)

# Save results
anim_save("animacaoo_casos_covidma.gif", animation = last_animation())
```

Finally, let's do the same for death rates from covid-19
```{r include = TRUE, warning = FALSE, message = FALSE}
# Save results
# Make the base map
map_ma_death <- ggplot() +
  geom_sf(data = ma_1)

# Plot animation
animation_ma <- map_ma_death +
  geom_sf(data = covid_ma,
          aes(size = last_available_deaths),
          colour = "red", alpha = .5) +
  scale_size_continuous(breaks = c(1,10,50,100,500,1000)) + 
  labs(size = "Confirmed cases") +
  ggtitle("Evolution of covid-19 deaths by municipality in Maranhão",
          subtitle = "Day {frame} of {nframes} since the first death") +
  labs(caption = "Source: brasil.io") +
  theme_minimal() +
  transition_manual(date)

# Setting speed of frames
animate(animation_ma, fps = 5)

# Save results
anim_save("animacaoo_obitos_covidma.gif", animation = last_animation())
```
